---
title: "Magic: The Gathering - A History of Stats"
author: "Tyson Franke"
date: "2026-03-26"
output: 
  html_document:
    theme:  
      bootswatch: "simplex"
      bg: "#f8f5f0"
      fg: "#222222"
      font_scale: 0.9
    highlight: "tango"
    toc: true
    toc_float: true
    css: "styles.css"
---

```{r setup, include=FALSE}
library(tidyverse)
library(jsonlite)
```

\{Introduction\}

## Dataset

We're using the default bulk dataset from scryfall. The dataset can be found here <https://scryfall.com/docs/api/bulk-data> or downloaded directly by clicking this link <https://data.scryfall.io/default-cards/default-cards-20260326090745.json>. The default bulk dataset contains only the English version of a printing, unless the card is only available in one language, e.g. printings exclusive to the set *Chronicles Foreign Black Border* or a Secret Lair card printed in the Phyrexian language.

**Note:** If downloading directly from the Scryfall website, the dataset will differ as it is updated frequently. For reproducibility, use the second, direct download link above.

In order to load this json file, the script uses jq to filter the data columns prior to loading in R for efficiency due to the large size. To install jq for Windows you can run `winget install jqlang.jq` in the terminal. Otherwise, for Mac/Linux and Windows alternatives refer to <https://jqlang.org/download/>. The dataset is then cached as an rds file. If loading a new dataset you will need to delete the cached rds file.

*When running the script the first time, it may still take a few minutes to load*

```{r preprocess, include=FALSE}
# Load from cached if available, otherwise file_path
file_path = "./default-cards-20260326090745.json"
rds_cache_path = "./cached_cards.rds"

# ALL COLUMNS: 
# > data <- fromJSON(file_path)
# > names(data)
#  [1] "object"              "id"                  "oracle_id"           "multiverse_ids"      "mtgo_id"             "arena_id"           
#  [7] "tcgplayer_id"        "cardmarket_id"       "name"                "lang"                "released_at"         "uri"                
# [13] "scryfall_uri"        "layout"              "highres_image"       "image_status"        "image_uris"          "mana_cost"          
# [19] "cmc"                 "type_line"           "oracle_text"         "colors"              "color_identity"      "keywords"           
# [25] "produced_mana"       "legalities"          "games"               "reserved"            "game_changer"        "foil"               
# [31] "nonfoil"             "finishes"            "oversized"           "promo"               "reprint"             "variation"          
# [37] "set_id"              "set"                 "set_name"            "set_type"            "set_uri"             "set_search_uri"     
# [43] "scryfall_set_uri"    "rulings_uri"         "prints_search_uri"   "collector_number"    "digital"             "rarity"             
# [49] "card_back_id"        "artist"              "artist_ids"          "illustration_id"     "border_color"        "frame"              
# [55] "full_art"            "textless"            "booster"             "story_spotlight"     "prices"              "related_uris"       
# [61] "purchase_uris"       "mtgo_foil_id"        "power"               "toughness"           "flavor_text"         "edhrec_rank"        
# [67] "penny_rank"          "all_parts"           "promo_types"         "security_stamp"      "card_faces"          "preview"            
# [73] "watermark"           "frame_effects"       "loyalty"             "printed_name"        "resource_id"         "tcgplayer_etched_id"
# [79] "flavor_name"         "attraction_lights"   "color_indicator"     "printed_type_line"   "printed_text"        "variation_of"       
# [85] "life_modifier"       "hand_modifier"       "content_warning"     "defense"

# SELECTED COLUMNS
cols <-  c("object", "id", "oracle_id", "name", "lang", "released_at", "layout", 
           "mana_cost", "cmc", "type_line", "oracle_text", "colors", "color_identity", "keywords", "produced_mana", 
           "legalities", "games", "reserved", "game_changer", "foil", "nonfoil", "finishes", "oversized", "promo", "reprint", 
           "set", "set_name", "set_type", "collector_number", "digital", "rarity", "artist", "border_color", "frame", "full_art", 
           "textless", "booster", "power", "toughness", "flavor_text", "loyalty"
           )

if (!file.exists("cached_cards.rds")) {
  jq_filter <- paste0(".[] | {", paste(cols, collapse=", "), "}")
  system2("jq", 
          args = c("-c", shQuote(jq_filter), shQuote(file_path)), 
          stdout = "jq_filtered_cards.ndjson")
  data <- stream_in(file("jq_filtered_cards.ndjson"), verbose = TRUE)
  saveRDS(data, rds_cache_path)
} else {
  print("Loading from cached dataset: cached_cards.rds")
  data <- readRDS(rds_cache_path)
}
```

```{r, echo=FALSE, class.output="glimpse-trunc"}
# Remaining columns after filtering via jq
glimpse(data)
```

## Data Cleaning

```{r cleaning}
df <- data %>%  filter(
    digital == FALSE,         # Only physical print (not MTGO or Arena exclusive)
    layout != "art_series"    # Omit art cards
  ) %>% 
  mutate(
    released_at = as.Date(released_at),
    year = year(released_at)
  )
```
