---
title: "mtg-anlaysis"
author: "Tyson Franke"
date: "2026-03-26"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(jsonlite)
```

## Dataset

We're using the default bulk dataset from scryfall. The dataset can be found here <https://scryfall.com/docs/api/bulk-data> or downloaded directly by clicking this link <https://data.scryfall.io/default-cards/default-cards-20260326090745.json>.

**Note:** If downloading directly from the Scryfall website, the dataset will differ as it is updated frequently. For reproducibility, use the second, direct download link above.

In order to load this json file, the script uses jq to filter the data prior to loading in R for efficiency due to the large size. To install jq for Windows you can run `winget install jqlang.jq` in the terminal. Otherwise, for Mac/Linux and Windows alternatives refer to <https://jqlang.org/download/>. The dataset is then cached as an rds file. If loading a new dataset you will need to delete the cached rds file.

*When running the script the first time, it may still take a few minutes to load*

```{r preprocess, include=FALSE}
# Load from cached if available, otherwise file_path
file_path = "./default-cards-20260326090745.json"
rds_cache_path = "./cached_cards.rds"

# ALL COLUMNS: 
# > data <- fromJSON(file_path)
# > names(data)
#  [1] "object"              "id"                  "oracle_id"           "multiverse_ids"      "mtgo_id"             "arena_id"           
#  [7] "tcgplayer_id"        "cardmarket_id"       "name"                "lang"                "released_at"         "uri"                
# [13] "scryfall_uri"        "layout"              "highres_image"       "image_status"        "image_uris"          "mana_cost"          
# [19] "cmc"                 "type_line"           "oracle_text"         "colors"              "color_identity"      "keywords"           
# [25] "produced_mana"       "legalities"          "games"               "reserved"            "game_changer"        "foil"               
# [31] "nonfoil"             "finishes"            "oversized"           "promo"               "reprint"             "variation"          
# [37] "set_id"              "set"                 "set_name"            "set_type"            "set_uri"             "set_search_uri"     
# [43] "scryfall_set_uri"    "rulings_uri"         "prints_search_uri"   "collector_number"    "digital"             "rarity"             
# [49] "card_back_id"        "artist"              "artist_ids"          "illustration_id"     "border_color"        "frame"              
# [55] "full_art"            "textless"            "booster"             "story_spotlight"     "prices"              "related_uris"       
# [61] "purchase_uris"       "mtgo_foil_id"        "power"               "toughness"           "flavor_text"         "edhrec_rank"        
# [67] "penny_rank"          "all_parts"           "promo_types"         "security_stamp"      "card_faces"          "preview"            
# [73] "watermark"           "frame_effects"       "loyalty"             "printed_name"        "resource_id"         "tcgplayer_etched_id"
# [79] "flavor_name"         "attraction_lights"   "color_indicator"     "printed_type_line"   "printed_text"        "variation_of"       
# [85] "life_modifier"       "hand_modifier"       "content_warning"     "defense"

# SELECTED COLUMNS
cols <-  c("object", "id", "oracle_id", "name", "lang", "released_at", "layout", 
           "mana_cost", "cmc", "type_line", "oracle_text", "colors", "color_identity", "keywords", "produced_mana", 
           "legalities", "games", "reserved", "game_changer", "foil", "nonfoil", "finishes", "oversized", "promo", "reprint", 
           "set", "set_name", "set_type", "collector_number", "digital", "rarity", "artist", "border_color", "frame", "full_art", 
           "textless", "booster", "power", "toughness", "flavor_text", "loyalty"
           )

if (!file.exists("cached_cards.rds")) {
  jq_filter <- paste0(".[] | {", paste(cols, collapse=", "), "}")
  system2("jq", 
          args = c("-c", shQuote(jq_filter), shQuote(file_path)), 
          stdout = "jq_filtered_cards.ndjson")
  data <- stream_in(file("jq_filtered_cards.ndjson"), verbose = TRUE)
  saveRDS(data, rds_cache_path)
} else {
  print("Loading from cached dataset: cached_cards.rds")
  data <- readRDS(rds_cache_path)
}
```


```{r}
head(data)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
